### 2025-03-27 sports-cognitive-psychology.qmd
20250327 - Yerkes-Dodson law

>>>>

The Yerkes–Dodson law describes the association between arousal and performance, as depicted in @fig-yerkesDodsonLaw.
For simple tasks, performance tends to increase with arousal.
For difficult tasks, there is an inverted-U-shaped association between arousal and performance, as depicted in @fig-hebbianYerkesDodsonLaw.
For difficult tasks, too little arousal is bad because it leads to insufficient attention on the task at hand; however, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking.
That is, there is an optimal range of arousal for challenging tasks, such as those in sports.

::: {#fig-yerkesDodsonLaw}
![](images/YerkesDodson.png){width=50% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg)"}

Yerkes–Dodson law. For simple tasks, performance increases with arousal. For difficult tasks, there is an inverted-U-shaped association between arousal and performance such that too little and too much arousal impairs performance. (Figure from [https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg](https://en.wikipedia.org/wiki/File:OriginalYerkesDodson.svg).)
:::

::: {#fig-hebbianYerkesDodsonLaw}
![](images/HebbianYerkesDodson.png){width=50% fig-alt="Yerkes–Dodson law. (Figure retrieved from https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg)"}

Yerkes–Dodson law for difficult tasks. For difficult tasks, there is an inverted-U-shaped association between arousal and performance. Too little arousal is bad because it leads to insufficient attention on the task at hand; by contrast, too much arousal is also bad because anxiety can impair performance through its negative effects on attention (e.g., "tunnel vision"), memory, decision making, problem-solving, and multitasking. (Figure from [https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg](https://en.wikipedia.org/wiki/File:HebbianYerkesDodson.svg).)
:::
